Top-Down Graph-Based Neuron Reconstruction

### 2017-11-13 papers/cvpr2018/0_introduction.tex
added HP comments

\section{Introduction}

The field of connectomics is concerned with reconstructing the wiring diagram of the brain at nanometer resolutions. 
Recent advancements in image acquisition using multi-beam serial-section electron microscopy (sSEM) has allowed neuroscientists to produce terabytes of electron micrsocopy (EM) image data every hour~\cite{hildebrand2017whole}.
Neuroscientists believe that reconstructing an entire mammalian brain at fine resolution will enable new insights into the workings of the brain~\cite{kasthuri2015saturated}. 
These observations will allow for new advancements in neuromedicine and artificial intelligence.

Segmenting the image stack into a label volume is a substantial component of this reconstruction process.
In a label volume, every voxel receives a segment label corresponding to a unique neuron. 
Voxels with the same label belong to the same neuron.
It is not feasible for domain experts to manually segment the vast amount of 3-D image data to model an entire brain.
Here we introduce a scalable top-down segmentation algorithm that also leverages local information.
The top-down partitioning strategy allows us to apply domain-specific constraints to the segmentation task.

A significant amount of research focuses on automatic reconstruction of the neurons in EM images~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image}. %TODO DOUBLE CHECK REFERENCES
All of these algorithms extract neuronal processes through the 3-D volumes using the raw image data as input.
Oftentimes, convolutional neural networks (CNNs) predict membrane probabilities or affinities between voxels and apply simple thresholds to agglomerate the voxels into clusters~\cite{lee2015recursive,ronneberger2015u}.
These \textit{pixel-based} algorithms produce excellent results but currently fall short of complete reconstructions with error rates of approximately $X\%$. %TODO ADD ERROR RATES

\begin{figure*}[h!]
	\centering
	\includegraphics[width=0.22\linewidth]{./figures/pipeline/image.png}
	\hspace{0.025\linewidth}
	\includegraphics[width=0.22\linewidth]{./figures/pipeline/affinities.png}
	\hspace{0.025\linewidth}
	\includegraphics[width=0.22\linewidth]{./figures/pipeline/watershed.png}
	\hspace{0.025\linewidth}
	\includegraphics[width=0.22\linewidth]{./figures/pipeline/neuroproof.png}
	\caption{An overview of an existing pipeline in segmentation for connectomics. From left to right: the original EM image data, the output of a CNN predicting voxel affinities, clustering of the affinities using a watershed algorithm, an agglomeration of the supervoxels into larger segments.}
	\label{fig:pipeline}
\end{figure*}

Researchers currently address the failures of these pixel-based algorithms by training random-forest classifiers to agglomerate an oversegmentation of voxels~\cite{nunez2014graph,10.1371/journal.pone.0125825}.
These classifiers take the output of the pixel-based algorithms as input and generate high-level statistics such as affinity distributions between pixel regions. 
Presently, these methods use hand-designed features despite the evidence that machine-learned features perform better~\cite{bogovic2013learned}. 
These \textit{region-based} algorithms outperform the pixel-based algorithms but still do not provide the accuracy needed for large scale reconstructions of the brain.

We present a \textit{segment-based} strategy that extends on the outputs of region-based methods.
We only take as input a segmentation from a previous agglomeration algorithm and not the raw image data. 
This allows us to train on connectomics datasets at a resolution and image type different than the test datasets.
We consider the 3-D skeletons of each segments as a simplification of its shape. 
We can then quickly extract a graph from the input segmentation using the skeletonization of the segments in the volume.

We add edge weights to our graph using a 3-D CNN to predict the probabilities that nearby segments belong to the same neuron.
Using a graph-based optimization strategy enables us to enforce domain-specific global constraints such as the underlying neuronal biology.
From here we apply a global partitioning algorithm that minimizes a cost function of separating adjacent nodes in the graph.

Our method takes the output of any bottom-up segmentation strategy and creates a graph on top of this segmentation. 
We enforce domain-specific global constraints by reasoning at the graph level.
This strategy scales to massive connectomics datasets and improves accuracy over existing methods.

>>>>

% !TEX root =  paper.tex
\section{Introduction}

The field of connectomics is concerned with reconstructing the wiring diagram of the brain at nanometer resolutions to enable new insights into the workings of the brain~\cite{kasthuri2015saturated,haehn2017scalable}. Recent advancements in image acquisition using multi-beam serial-section electron microscopy (sSEM) have allowed researchers to produce terabytes of image data every hour~\cite{hildebrand2017whole}. It is not feasible for domain experts to manually reconstruct thousands or millions of neurons from this vast amount of image data~\cite{haehn2014design}. State-of-the-art automatic reconstruction approaches use pixel-based segmentation with convolutional neural networks (CNNs) followed by agglomeration strategies~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,lee2015recursive,ronneberger2015u}. %TODO DOUBLE CHECK REFERENCES
These \textit{bottom-up pixel-based} algorithms produce excellent results but still fall short of acceptable error rates for large volumes.

\begin{figure*}[t!]
	\centering
	(a) \includegraphics[width=0.18\linewidth]{./figures/pipeline/image.png}
	\hspace{0.025\linewidth}
	(b) \includegraphics[width=0.18\linewidth]{./figures/pipeline/affinities.png}
	\hspace{0.025\linewidth}
	(c) \includegraphics[width=0.18\linewidth]{./figures/pipeline/watershed.png}
	\hspace{0.025\linewidth}
	(d) \includegraphics[width=0.18\linewidth]{./figures/pipeline/neuroproof.png}
	\caption{An overview of a pixel-based connectomics segmentation pipeline. (a) Original EM image data (b) Output of a CNN predicting voxel affinities (c) Clustering of the affinities using a watershed algorithm (d) Agglomeration of the super-voxels into larger segments.}
	\label{fig:pipeline}
\end{figure*}

%Researchers address the failures of these pixel-based algorithms by training random-forest classifiers to agglomerate an oversegmentation of voxels~\cite{nunez2014graph,10.1371/journal.pone.0125825}.
%These classifiers take the output of the pixel-based algorithms as input and generate high-level statistics such as affinity distributions between regions.
%Presently, these methods use hand-designed features despite the evidence that machine-learned features perform better~\cite{bogovic2013learned}.
%These \textit{region-based} algorithms outperform the pixel-based algorithms but do not provide the accuracy needed for large scale reconstructions of the brain.

%Here we introduce a scalable, top-down segmentation algorithm that also leverages local information.
%Partitioning globally allows us to apply domain-specific constraints to the segmentation task.
%We extract a graph representation from the input segmentation.
%By simplifying the 3-D segments by their skeletons, we can quickly generate the graph nodes and edges.

We present a \textit{top-down graph-based} method that builds on the outputs of bottom-up pixel-based segmentation approaches (Fig.~\ref{fig:teaser}a). We first extract 3D skeletons from the input segmentation (Fig.~\ref{fig:teaser}b) and generate a simplified 3D graph (Fig.~\ref{fig:teaser}c). We train a 3D CNN classifier on the agglomerated regions in the segmentation data to detect errors. During test time, we run the classifier to populate the graph edge weights with error probabilities. We then use a graph optimization algorithm to partition the graph into the final reconstruction by enforcing domain-specific global constraints from the underlying biology (Fig.~\ref{fig:teaser}d).

Our approach operates at a level of abstraction above existing pixel-based methods. This allows us to leverage both local and global information to produce more accurate reconstructions. Because it uses agglomerated regions our classifier is independent of image resolution and acquisition parameters, enabling its application to isotropic and anisotropic image data without retraining. Using the 3D graph of the segmentation allows us to enforce global constraints on the reconstruction that closely follow the underlying biology, such as limiting the angles of neuron branchings to the biologically plausible range. We can also use local information such as shape priors to weight the edges in our graph. This dual approach of assessing local decisions in a global context yields accuracy improvements over the existing reconstruction methods.

This work makes the following following contributions: (1) a novel top-down graph-based method for neural reconstruction of connectomics data; (2) a region-based CNN classifier to detect errors under global constraints; (3) an empirical evaluation of our method on several connectomics datasets; (4) our method yields improved performance over the state-of-the-art on X out of Y datasets, and competitive results on the remaining data. On average, we improve the Variation of Information (VI) metric across all datasets by X percent without drastically increasing the running time.

### 2017-11-13 papers/cvpr2018/1_related_work.tex
added HP comments

\section{Related Work}

A significant amount of research in computer vision focuses on the segmentation of images \cite{zaitoun2015survey}. 
Here, we review some of the most successful methods that have been applied to large-scale EM images. Figure \ref{fig:pipeline} shows the results of a typical segmentation pipeline on these images.

\paragraph{Pixel-based methods.}

A large amount of connectomics research considers the problem of extracting segmentation information at the voxel level from the raw EM images to produce label volumes.
Some early techniques apply computationally expensive graph partitioning algorithms with a single node per voxel~\cite{andres2012globally}, however these methods do not scale to terabyte datasets.
Since then, several methods train 2-D CNNs to predict membrane probabilities per image slice~\cite{ciresan2012deep,jain2010boundary,kaynig2015large,rhoananet,amelio_segmentation}. 
More recent strategies augment these neural networks with the $z$-dimension creating full 3-D CNNs~\cite{lee2015recursive,ronneberger2015u}.
Oftentimes these networks produce probabilities for the affinity between two voxels (i.e., the probability that adjacent voxels belong to the same neuron).
The MALIS cost function is specifically designed for generating affinities that produce good segmentations~\cite{briggman2009maximin}. 
More recently, flood-filling networks produce segmentations by training an end-to-end neural network that takes EM images to label volumes~\cite{januszewski2016flood}.
These flood-filling networks produce very impressive accuracies but at too high a computational cost.

\paragraph{Region-based methods.}

Several of these pixel-based approaches generate probabilities that neighboring voxels belong to the same neuron.
Many algorithms build on top of these methods and train random-forest classifiers to produce segmentations of the EM images~\cite{seymour2016rhoananet,nunez2014graph,10.1371/journal.pone.0125825,parag2017anisotropic,zlateski2015image}. 
Despite the success of these classifiers, they often make mistakes based on the local information leading to errors in the segmentation. 
Lately, many recent advancements in segmentation use artificial neural networks since machine-learned features outperform hand-designed ones~\cite{bogovic2013learned}.
A significant amount of research focuses on the training and optimization of these deep neural networks~\cite{chatfield2014return,maas2013rectifier,nesterov1983method}.

\paragraph{Segment-based methods.}

Some limited research builds on top of the region-based methods to correct errors in the segmentation \cite{rolnick2017morphological,error_correction_using_CNN,haehn2017guided}. 
However, to our knowledge, we are the first to extract a graph representation from these region-based methods. 
Many segmentation and clustering algorithms use graph partitioning techniques~\cite{andres2012globally} or normalized cuts for traditional image segmentation~\cite{kappes2016higher,shi2000normalized,tatiraju2008image}. 
We can enforce domain-specific constraints using graph-based paritioning algorithms.  
Unfortunately graph partitioning is an NP-Hard problem~\cite{demaine2006correlation}.
However, there are several useful multicut heuristics which provide good approximations with reasonable computational costs~\cite{horvnakova2017analysis,keuper2015efficient}.

>>>>

% !TEX root =  paper.tex
\section{Related Work}

A significant amount of research in computer vision focuses on the segmentation of images~\cite{zaitoun2015survey}. Here, we review some of the most successful methods that have been applied to large-scale EM images in connectomics. Fig.~\ref{fig:pipeline} shows the results of a typical connectomics segmentation pipeline.

%\paragraph{Pixel-based methods.}

A large amount of connectomics research considers the problem of extracting segmentation information at the pixel (i.e., voxel) level from the raw EM images. Some early techniques apply computationally expensive graph partitioning algorithms with a single node per pixel~\cite{andres2012globally}. However these methods do not scale to terabyte datasets. More recent methods train classifiers to predict membrane probabilities per image slice either using 2D~\cite{ciresan2012deep,jain2010boundary,kaynig2015large,seymour2016rhoananet,amelio_segmentation} or 3D CNNs~\cite{lee2015recursive,ronneberger2015u}.\hp{add more refs}

Oftentimes these networks produce probabilities for the affinity between two voxels (i.e., the probability that adjacent voxels belong to the same neuron). The MALIS cost function is specifically designed for generating affinities that produce good segmentations~\cite{briggman2009maximin}. \hp{add seung ref} More recently, flood-filling networks produce segmentations by training an end-to-end neural network that goes from EM images directly to label volumes~\cite{januszewski2016flood}. These flood-filling networks produce impressive accuracies but at a high computational cost.

%\paragraph{Region-based methods.}

%Lately, many recent advancements in segmentation use artificial neural networks since machine-learned features outperform hand-designed ones~\cite{bogovic2013learned}.
%A significant amount of research focuses on the training and optimization of these deep neural networks~\cite{chatfield2014return,maas2013rectifier,nesterov1983method}.

Several pixel-based approaches generate probabilities that neighboring pixels belong to the same neuron.
Many methods build on top of these methods and train random-forest classifiers \hp{really?} to produce region-based (i.e., super-pixel) segmentations~\cite{seymour2016rhoananet,nunez2014graph,10.1371/journal.pone.0125825,parag2017anisotropic,zlateski2015image}. \hp{are these agglomeration techniques based on watershed?}

%\paragraph{Segment-based methods.}

Some recent research builds on top of these region-based methods to correct errors in the segmentation~\cite{rolnick2017morphological,error_correction_using_CNN,haehn2017guided}.
However, to our knowledge, our method is the first to extract a 3D graph from pixel-based agglomerated segmentations for a true top-down reconstruction approach. This allows us to enforce domain-specific constraints using graph-based partitioning algorithms. Many segmentation and clustering algorithms use graph partitioning techniques~\cite{andres2012globally} or normalized cuts for traditional image segmentation~\cite{kappes2016higher,shi2000normalized,tatiraju2008image}.
Even though graph partitioning is an NP-Hard problem~\cite{demaine2006correlation} there are several useful multicut heuristics that provide good approximations with reasonable computational costs~\cite{horvnakova2017analysis,keuper2015efficient}. We use the method of X et al.\hp{add ref} to partition the extracted 3D graph into the final neuron reconstruction.

### 2017-11-13 papers/cvpr2018/2_method.tex
added HP comments

There are two types of errors that can occur in image segmentation problems.
The first, called a split error, occurs when there are two segments that should have been merged.
The second, called a merge error, happens when one segment should be split into two.
Generally, it is much more difficult to correct merge errors than to correct split errors (CITE). 
Our method takes as input an oversegmentation of an $EM$ image volume where there are many more split errors than merge errors.
Our goal is to identify locations of split errors and merge the corresponding segments.
Section~\ref{sec:neuroproof} discusses two pipelines for generating these segmentations.

From the input segmentation we generate a graph $G$ with nodes $N$ and edges $E$ with weights $w_e$. 
The nodes correspond to label segments from the segmentation with edges between segments considered for merging.
Ideally, our graph has edges corresponding to all of the segments which were erroneously split with few edges between correctly split segments.
We generate a skeleton for every segment following the intuition that a skeleton represents a simplified representation of the overall shape of an object. 
We locate merge candidates and produce corresponding edges based on these skeletons.
A CNN generates the edge weights by learning a merge function given two segments as input. 
A multicut heuristic generates a partition on the graph where nodes in the same partition are assigned the same output label. 
Thus there are three major components to our framework: skeletonization for constructing a graph, network training for determining edge weights, and graph-partitioning using a multicut heuristic algorithm.

>>>>

There are two types of errors that can occur in connectomics segmentation.
The first, called a split error, occurs when there are two segments that should have been merged. The second, called a merge error, happens when one segment should be split into two. Generally, it is much more difficult to correct merge errors than to correct split errors (CITE)\hp{add}, which is why most reconstruction approaches are tuned towards over-segmentation with many more split than merge errors. Our method takes as input over-segmentations of EM image volumes generated by state-of-the-art connectomics reconstruction pieplines (Sec.~\ref{sec:neuroproof}). Our goal is to identify locations of split errors and merge the corresponding segments automatically.

From the input segmentation we generate a graph $G$ with nodes $N$ and edges $E$ with weights $w_e$. The nodes correspond to label segments from the segmentation with edges between segments considered for merging. Ideally, our graph has edges corresponding to all of the segments which were erroneously split with few edges between correctly split segments. We generate a skeleton for every segment based on the intuition that a skeleton represents a simplified representation of the overall shape of the neuron. We generate a simplified 3D graph from the skeletons by locating merge candidates and producing corresponding edges.\hp{check prior sentence} We train a CNN to  generate the edge weights by learning a merge function from ground truth data. A multicut heuristic generates a partition on the graph where nodes in the same partition are assigned the same output label. We will now discuss the three major components to our framework (skeletonization, CNN, and graph-partition) in more detail. \hp{stopped here}

### 2017-11-13 papers/cvpr2018/paper.tex
added HP comments

@@

%%%%%%%%% TITLE
\title{Top-down Segment-based Neuron Reconstruction}

@@

\begin{figure}[t]\footnotesize
	\vspace{-2.04in}
	\begin{minipage}{\textwidth}
		\begin{minipage}{0.125\textheight}
			\includegraphics[width=\linewidth]{figures/schema/teaser-image.png}
			\includegraphics[width=\linewidth]{figures/schema/teaser-segmentation.png}
		\end{minipage}
		\begin{minipage}{0.24\textwidth}
			\includegraphics[width=\linewidth]{figures/schema/pre-multicut.png}
		\end{minipage}
		\begin{minipage}{0.34\textwidth}
			\includegraphics[width=\linewidth]{figures/schema/multicut-graph.png}
		\end{minipage}
		\begin{minipage}{0.24\textwidth}
			\includegraphics[width=\linewidth]{figures/schema/post-multicut.png}
		\end{minipage}
		\caption{To reconstruct neurons, we take as input the results of a region-based agglomeration algorithm. We extract a graph from this segmentation and apply a partitioning algorithm with local and global features to improve accuracy.}
		\label{fig:teaser}
	\end{minipage}
\end{figure}

%%%%%%%%% ABSTRACT
\begin{abstract}
Advancements in electron microscopy image acquisition have created massive connectomics datasets which are petabytes in size and make manual segmentation infeasible. 
Proposed methods for automatic segmentations generate super-pixel approximations of neural membranes followed by agglomeration strategies to create full neuron reconstructions. 
However, pixel-based segmentation results widely neglect global geometric properties.
We generate skeletons of membrane labels to approximate neural pathways in 3D. This allows us to efficiently apply graph-based optimization strategies and to train automatic classifiers for shape description against manually labeled ground truth. Our classifiers detect feasible and impossible neural pathways by looking at geometric properties alone, and are able to improve the segmentation labels accordingly. We demonstrate the performance of our classifiers on multiple real-world connectomics datasets with average variation of information improvement of $X\times$. We discuss our findings and provide insights to learning shape features for neuron agglomeration.
\end{abstract}

>>>>

\newcommand{\hp}[1]{\textcolor{red}{HP: #1}}

@@

%%%%%%%%% TITLE
\title{Top-Down Graph-Based Neuron Reconstruction}

@@

\begin{figure}[t]\footnotesize
	\vspace{-2.3in}
	\begin{minipage}{\textwidth}
		\begin{minipage}{0.2\textwidth}
			 \includegraphics[width=\linewidth]{figures/schema/teaser-image.png}
			(a)\includegraphics[width=\linewidth]{figures/schema/teaser-segmentation.png}
		\end{minipage}
		\begin{minipage}{0.2\textwidth}
			(b)\includegraphics[width=\linewidth]{figures/schema/pre-multicut.png}
		\end{minipage}
		\begin{minipage}{0.3\textwidth}
			(c)\includegraphics[width=\linewidth]{figures/schema/multicut-graph.png}
		\end{minipage}
		\begin{minipage}{0.2\textwidth}
			(d)\includegraphics[width=\linewidth]{figures/schema/post-multicut.png}
		\end{minipage}
		\caption{Outline of our approach. (a) Raw EM image data (top) and results of pixel-based segmentation and agglomeration algorithms (bottom). (b) Extracted 3D skeleton from the segmentation. (c) Simplified 3D graph with error probabilities from trained classifier (schematic). (d) Result of graph partitioning algorithm with local and global geometric features (schematic). (d) Improved 3D reconstruction.}
		\label{fig:teaser}
	\end{minipage}
\end{figure}

%%%%%%%%% ABSTRACT
\begin{abstract}
Advancements in electron microscopy image acquisition have created massive connectomics datasets in the terabyte range that make manual reconstruction of neuronal structures infeasible. Current state-of-the-art automatic methods segment neural membranes at the pixel level followed by agglomeration methods to create full neuron reconstructions. However, these approaches widely neglect global geometric properties that are inherent in the graph structure of neural wiring diagrams. Instead, we follow bottom-up pixel-based reconstruction by a top-down graph-based method to more accurately approximate neural pathways. Using the membrane labels of the pixel-based segmentation we first generate skeletons in 3D. Using this 3D graph we train automatic classifiers for shape description to detect impossible neural pathways by looking at geometric properties. We then apply efficient graph-based optimization strategies to improve the segmentation labels. We demonstrate the performance of our approach on multiple real-world connectomics datasets with average variation of information improvement of $X\times$. The paper provides insights into learning shape features for top-down graph-based neuron reconstruction.
\end{abstract}
